Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-07-05 survey/survey.tex
Finish section 4.  I might add a reference or two, but the core content is, AFAIK, all there.

\newcommand{\idd}{\mathit{id}\leftrightarrow}
\newcommand{\idf}{\mathit{id}\Leftrightarrow}

\begin{tikzcd}[column sep=normal, row sep=small]
 && (a+c)+b \arrow [r, "\swapp \oplus\idd", ""{name=U, below}] & (c+a)+b \arrow [dr, "\assocrp"] && \\
 & a+(c+b) \arrow [ur, "\assoclp"] & & & c+(a+b) \arrow [dr, "\idd\oplus\swapp"] &  \\
a+(b+c) \arrow [ur, "\idd\oplus\swapp"] \arrow [r, "\assoclp"]
  \arrow [dr, "\assoclp"]
  \arrow [ddr, swap, "\assoclp"]
    & (a+b)+c \arrow [r, "\swapp"] &
    c+(a+b) \arrow [r, swap, "\assoclp", ""{name=D, above}] 
    & |[alias=Z]| (c+a)+b \arrow [r, "\assocrp"] &c+(a+b) \arrow [r, "\idd\oplus\swapp"] & c+(b+a) \\
 & (a+b)+c \arrow [dr, "\swapp"] &&&& \\
 & (a+b)+c \arrow [dr, swap, "\swapp"] & c+(a+b) \arrow [rr, swap, "\idd", ""{name=DD, above}]
             \arrow [d, Rightarrow, "\idf\, \mathit{idl}\odot{l}"] && 
    c+(a+b) \arrow [ruu, "\idd\oplus\swapp"] & \\
 && c+(a+b) \arrow [rrruuu, bend right = 30, swap, "\idd\oplus\swapp"] && \\
 \arrow[Rightarrow, from=U, to=D, "\mathit{hexagon}\oplus{r}\, \boxdot\, \idf"]
 \arrow[Rightarrow, from=Z, to=DD, swap, "\idf\boxdot\mathit{linv}\odot{l}\,\boxdot\,\idf"]
\end{tikzcd}

@@

We would be remiss in letting the reader believe that the above is ``the''
categorical diagram that would be found in categorical textbooks. Rather,
congruence would be used to elide the $\idf$. The various arrows would also
be named differently --- our \assoclp is often named $\alpha$,
\assocrp is $\alpha^{-1}$, $\swapp$ is $B$ (always with subscripts).

>>>>

\newcommand{\idd}{\mathit{id}\leftrightarrow}
\newcommand{\idf}{\mathit{id}\Leftrightarrow}
\vspace*{3mm}
\begin{tikzcd}[column sep=normal, row sep=normal]
 && (a+c)+b \arrow [r, "\swapp \oplus\idd", ""{name=U, below}] & (c+a)+b \arrow [dr, "\assocrp"] && \\
 & a+(c+b) \arrow [ur, "\assoclp"] & & & c+(a+b) \arrow [dr, "\idd\oplus\swapp"] &  \\
a+(b+c) \arrow [ur, "\idd\oplus\swapp"] \arrow [r, "\assoclp"]
  \arrow [dr, "\assoclp"]
  \arrow [ddr, swap, "\assoclp"]
    & (a+b)+c \arrow [r, "\swapp"] &
    c+(a+b) \arrow [r, swap, "\assoclp", ""{name=D, above}] 
    & |[alias=Z]| (c+a)+b \arrow [r, "\assocrp"] &c+(a+b) \arrow [r, "\idd\oplus\swapp"] & c+(b+a) \\
 & (a+b)+c \arrow [dr, "\swapp"] &&&& \\
 & (a+b)+c \arrow [dr, swap, "\swapp"] & c+(a+b) \arrow [rr, swap, "\idd", ""{name=DD, above}]
             \arrow [d, Rightarrow, "\idf\, \mathit{idl}\odot{l}"] && 
    c+(a+b) \arrow [ruu, "\idd\oplus\swapp"] & \\
 && c+(a+b) \arrow [rrruuu, bend right = 40, swap, "\idd\oplus\swapp"] && \\
 \arrow[Rightarrow, from=U, to=D, "\mathit{hexagon}\oplus{r}\, \boxdot\, \idf"]
 \arrow[Rightarrow, from=Z, to=DD, swap, "\idf\boxdot\mathit{linv}\odot{l}\,\boxdot\,\idf"]
\end{tikzcd}

@@

We would be remiss in letting the reader believe that the above is ``the''
categorical diagram that would be found in categorical textbooks. Rather,
congruence would be used to elide the $\idf$. Furthermore, the various arrows
would also be named differently --- our \assoclp is often named $\alpha$,
\assocrp is $\alpha^{-1}$, $\swapp$ is $B$ (always with subscripts).
And the two steps needed to remove inverses (i.e. first cancelling
inverse arrows, then removing the resulting identity arrow ``in context'')
are often combined into one. Here we'll simply name this operation 
$\mathit{cancel}$, which could be programmed as a defined function over
$\Pi$ level-2.  The result would then be the much simpler

\vspace*{3mm}
\begin{tikzcd}[column sep=normal, row sep=normal]
 & a+(c+b) \arrow [r, "\assoclp", ""{name=U, below}] & (a+c)+b \arrow [rd, "\swapp\oplus\idd"] & & & \\
a+(b+c) \arrow [ur, "\idd\oplus\swapp"] \arrow [dr, "\assoclp"]
  & & & (c+a)+b \arrow [r, "\assocrp", ""{name=UU, below}] & c+(a+b) \arrow [r, "\idd\oplus\swapp"] & c+(b+a) \\
 & (a+b)+c \arrow [r, swap, "\swapp", ""{name=D, above}] & c+(a+b) \arrow [ur, "\assoclp"] 
  \arrow [urrr, swap, "\idd\oplus\swapp", ""{name=DD,above}] & & & & \\
 \arrow[Rightarrow, from=U, to=D, "\mathit{hexagon}\oplus{r}"]
 \arrow[Rightarrow, from=UU, to=DD, "\mathit{cancel}"]
\end{tikzcd}

In other words, each (non-refocusing) line of the proof of
\AgdaFunction{swap{-}fl2⇔swap{-}fl1}\; is a complete path 
from left to right in each diagram above, and the annotation
on the right-hand-side becomes the natural transformation (denoted
by vertical $\Rightarrow$) justifying the move to the next line.
The first diagram uses lines $1,4,7,8$ in full; the second
diagram collapses $7$ and $8$ into one, as well as not duplicating
parts which are related by $\idf$.
